---
title: "Geog533 Inclass 4"
author: "Dr. Qiusheng Wu"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

**Questions:** Snowfall for a location is found to be normally distributed with mean 96″ and standard deviation of 32″.

## Question 1a
**(a) What is the probability that a given year will have 100″ of snow?**
```{r}
answer_a <- dnorm(100,mean = 96,sd = 32)
print(answer_a)
```

The answer is **`r answer_a`**.

Plot the density function:
```{r}
curve(dnorm(x,mean = 96,32),from = 0,to = 192)
points(x = 100,y = answer_a,pch=16)
```

## Question 1b
**(b) What is the probability that a given year will have more than 120″ of snow?**

```{r}
answer_b <- pnorm(120,mean = 96,sd = 32,lower.tail = FALSE)
print(answer_b)
```
The answer is **`r answer_b`**.

Plot the density function
```{r}
curve(dnorm(x,mean = 96,32),from = 0,to = 192)
cord.x <- c(120,seq(120,192,by = 0.01),192)
cord.y <- c(0,dnorm(seq(120,192,0.01),mean = 96,sd = 32),0)
polygon(cord.x,cord.y,col="red")
```

## Question 1c
**(c) What is the probability that a given year will have less than 80″ of snow?**
```{r}
answer_c <- pnorm(80,mean = 96,sd = 32)
print(answer_c)
```
The answer is **`r answer_c`**

Plot the density function:
```{r}
curve(dnorm(x,mean = 96,32),from = 0,to = 192)
cord.x <- c(0,seq(0,80,by = 0.01),80)
cord.y <- c(0,dnorm(seq(0,80,0.01),mean = 96,sd = 32),0)
polygon(cord.x,cord.y,col="blue")
```

## Question 1d
**(d) What is the probability that the snowfall will be between 90″ and 100″?**
```{r}
answer_d <- diff(pnorm(c(90,100),mean = 96,sd = 32))
print(answer_d)
```
The answer is **`r answer_d`**

Plot the density function:
```{r}
curve(dnorm(x,mean = 96,32),from = 0,to = 192)
cord.x <- c(90,seq(90,100,by = 0.01),100)
cord.y <- c(0,dnorm(seq(90,100,0.01),mean = 96,sd = 32),0)
polygon(cord.x,cord.y,col="green")
```


## Question 1e
**(e) What level of snowfall will be exceeded only 10% of the time?**


```{r}
answer_e <- qnorm(0.1,mean = 96,sd = 32,lower.tail = FALSE)
print(answer_e)
```

The answer is **`r answer_e`**

Plot the density function:
```{r}
curve(dnorm(x,mean = 96,32),from = 0,to = 192)
cord.x <- c(137,seq(137,192,by = 0.01),192)
cord.y <- c(0,dnorm(seq(137,192,0.01),mean = 96,sd = 32),0)
polygon(cord.x,cord.y,col="black")
```

