---
title: "Geog533 Lab 8 - Bivariate Regression"
author: "Dr. Qiusheng Wu"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

Complete the following exercises in Chapter 8 (Introduction to Regression Analysis) of the textbook [R] pages 247-250. 

## Question 1

This is Exercise 1 in Chapter 8 of the Textbook [R].

A regression of weekly shopping trip frequency on annual income (data entered in thousands of dollars) is performed on data collected from 24 respondents. The results are summarized below:

* Intercept: 0.46
* Slope: 0.19

|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|                 |                    |                |                 |
| Residual |       1.7       |                    |                |                 |
| Total    |       2.3       |                    |                |                 |


### Question 1(a)
Fill in the blanks in the ANOVA table.

```{r}
q1.n <- 24
q1.intercept <- 0.46

q1.slope <- 0.19

q1.TSS <- 2.3  # Total sum of squares
q1.RSS <- 1.7  # Residual sum of squares
q1.MSS <- q1.TSS - q1.RSS # Model sum of squares

q1.df1 <- 1 
q1.df2 <- q1.n-2

q1.msq1 <- q1.MSS/q1.df1 # mean square of regression  0.6
q1.msq2 <- q1.RSS/q1.df2  # mean square of residual   0.077

q1.F.value <- q1.msq1/q1.msq2  # 7.76

```


|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|   `r q1.MSS`    |    `r q1.df1`      |    `r q1.msq1` |  `r q1.F.value` |
| Residual |       1.7       |    `r q1.df2`      |     `r q1.msq2`|                 |
| Total    |       2.3       |  `r q1.df1+q1.df2` |                |                 |


### Question 1(b)
What is the predicted number of weekly shopping trips for someone making $50,000/year?
```{r}
y <- q1.intercept + q1.slope * (50000/1000)  # y = 9.96
y
```


### Question 1(c)
In words, what is the meaning of the coefficient 0.19?

The meaning is that **for every increase in annual income equal to $1000, the predicted weekly shopping trip frequency will increase by 0.19.**

### Question 1(d)
Is the regression coefficient significantly different from zero? How do you know?
```{r}
q1.F.critical <- qf(0.95,q1.df1,q1.df2)  # 4.30
if(q1.F.value > q1.F.critical) {
  print("We reject the null hypothesis")
} else {
  print("We fail to reject the null hypothesis")
}

# since F.value > F.critical, we reject the null hypothesis. In other words, the regression coefficient is significantly different from zero

```



### Question 1(e)
What is the value of the correlation coefficient?
```{r}
q1.r.sq <- q1.MSS/q1.TSS  # R2 = 0.26
q1.r <- sqrt(q1.r.sq)  # r = 0.51
```

The correlation coefficient is `r round(q1.r,2)`



## Question 2

This is Exercise 6 in Chapter 8 of the Textbook [R].

The following data are collected in an effort to determine whether snowfall is dependent upon elevation: 

| Snowfall (inches) | Elevation (feet) |
|-------------------|------------------|
|       36          |       400        |
|       78          |       800        |
|       11          |       200        |
|       45          |       675        |

Using R, show your work on exercises (a) through (g).

### Question 2(a)
Find the regression coefficients (the intercept and the slope coefficient).
```{r}
y <- c(36,78,11,45)
x <- c(400,800,200,675)
n <- length(y)

m <- lm(y~x)
summary(m)

intercept <- coefficients(m)[1]
slope <- coefficients(m)[2]
```

**intercept = `r intercept`**

**slope = `r slope`**


### Question 2(b)
Estimate the standard error of the residuals about the regression line.
```{r}
# use the equation on page 236 of the textbook
se <- sqrt(sum(residuals(m)^2)/(n-2))  # 11.252
```

The standard error of the residuals is: **`r se`**

### Question 2(c)
Test the hypothesis that the regression coefficient associated with the independent variables is equal to zero. Also place a 95% confidence interval on the regression coefficient.
```{r}
summary(m)
anova(m)
# p-value: 0.05663
# p-value > 0.05, so we fail to reject the null hypothesis that the regression coefficient associated with the independent variables is equal to zero
confint(m)
#                     2.5 %     97.5 %
# (Intercept)   -66.487337152 51.2129252
# x             -0.006754067  0.2000541
```


### Question 2(d)
Find the value of $r^2$.
```{r}
summary(m)
# R2 = 0.8899
```


### Question 2(e)
Make a table of the observed values, predicted values, and the residuals.
```{r}
df <- data.frame(x,y)
df$predict <- fitted(m)
df$residual <- residuals(m)
library(knitr)
kable(df)
```


### Question 2(f)
Prepare an analysis of variance table portraying the regression results.
```{r}
am <- anova(m)
MSS <- am["Sum Sq"][[1]][1]
RSS <- am["Sum Sq"][[1]][2]
TSS <- MSS+RSS
df1 <- am["Df"][[1]][1]
df2 <- am["Df"][[1]][2]
df3 <- df1+df2
msq1 <- am["Mean Sq"][[1]][1]
msq2 <- am["Mean Sq"][[1]][2]
F.value <- am["F value"][[1]][1]
```


|          | Sum of squares  | Degrees of freedom |  Mean square   |       *F*       |
|----------|-----------------|--------------------|----------------|-----------------|
|Regression|    `r MSS`      |       `r df1`      | `r msq1`       |    `r F.value`  |
| Residual |    `r RSS`      |       `r df2`      |`r msq2`        |                 |
| Total    |    `r TSS`      |       `r df3`      |                |                 |

### Question 2(g)
Graph the data and the regression line.
```{r}
plot(x,y)
abline(m,col="red",lwd=2)
```




## Question 3

This is Exercise 10 in Chapter 8 of the Textbook [R].

Use R and the [Milwaukee dataset](http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv) to:

### Problem
perform a regression using sales price as the dependent variable and lot size as the independent variable.

### Solution
```{r}
url = "http://spatial.binghamton.edu/geog533/data/Milwaukee_Sales.csv"
df <- read.csv(url,header = TRUE)
m <- lm(SalePrice ~ LotSize,data = df)
summary(m)
plot(df$LotSize,df$SalePrice)
abline(m,col="red",lwd=2)

```



## Question 4

This is Exercise 11 in Chapter 8 of the Textbook [R].

Use R and the Hypothetical [UK Housing Prices dataset](http://spatial.binghamton.edu/geog533/data/UK_Housing.csv) to:

### Question 4(a) 
perform a regression using house price as the dependent variable, and number of bedrooms as the independent variable;

```{r}
url = "http://spatial.binghamton.edu/geog533/data/UK_Housing.csv"
df <- read.csv(url,header = TRUE)
m <- lm(df$price~df$bedrooms)
summary(m)
```


### Question 4(b) 
repeat part (a), using the number of bathrooms as the independent variable, and comment on the results.
```{r}
m <- lm(df$price~df$bathrooms)
summary(m)
```


## Question 5
Download the [hometown.csv](http://spatial.binghamton.edu/geog533/data/hometown.csv), append your name and hometown coordinates to the csv, then use [leaflet](https://rstudio.github.io/leaflet/) to create a web map (10 pt.)

```{r, message=FALSE, warning=FALSE}
if(!require(leaflet)) install.packages("leaflet")
library(leaflet)
df <- read.csv("http://spatial.binghamton.edu/leaflet/hometown.csv")
#df <- read.csv("hometown.csv")
df$latitude = as.character(lapply(strsplit(as.character(df$location), split=","), "[", 1))
df$longitude = as.character(lapply(strsplit(as.character(df$location), split=","), "[", 2))
df$fullname <- paste(df$firstname,df$lastname)  
df$popup <- paste("<center><h4>",df$fullname,"</h4></center>","<img src=",df$image," width='200px'/>",sep = "")
#write.csv(df,"hometown.csv",row.names = FALSE)
leafIcons <- icons(
  iconUrl = ifelse(df$type == "grads",
                   "http://spatial.binghamton.edu/leaflet/images/Blue.png",
                   "http://spatial.binghamton.edu/leaflet/images/Red.png"))
leaflet(df,width = "100%",height = 800) %>% 
  addTiles(group = "OSM (default)") %>%
  addProviderTiles(provider = "Esri.WorldStreetMap",group = "World StreetMap") %>%
  addProviderTiles(provider = "Esri.WorldImagery",group = "World Imagery") %>%
  addProviderTiles(provider = "NASAGIBS.ViirsEarthAtNight2012",group = "Nighttime Imagery") %>%
  addTiles() %>% fitBounds(147,90,-101,-90) %>% 
  addMarkers(lng = as.numeric(df$longitude), lat = as.numeric(df$latitude), icon =leafIcons,popup = df$popup,clusterOptions = markerClusterOptions())  %>% 
  addLayersControl(
    baseGroups = c("OSM (default)","World StreetMap", "World Imagery", "Nighttime Imagery"),
    options = layersControlOptions(collapsed = FALSE)
  )
```

