---
title: "Geog533 Lab 10"
author: "Qiusheng Wu"
output: 
  html_notebook:
    toc: TRUE
    toc_float: TRUE
---

## Question 1
Use R package UScensus2010county to complete the following tasks:  (20 pt.)

### Question 1(a)
Plot a map of New York counties using the plot function.
```{r, message=FALSE, warning=FALSE}
# if (!require(UScensus2010)) {install.packages("UScensus2010")}
# if(!require(UScensus2010county)) {install.county("osx")}
# if(!require(UScensus2000tract)) {install.tract("osx")}
library(UScensus2010)
library(UScensus2010county)
library(UScensus2010tract)

data("new_york.county10")
county <- new_york.county10
plot(county)
county.df <- county@data
county.df
```

### Question 1(b)	
Plot a map of New York counties using the qtm function.
```{r, message=FALSE, warning=FALSE}
if(!require(tmap)) {install.packages("tmap")}
library(tmap)
qtm(shp = county,fill = "NAME10",fill.title="County")
```


### Question 1(c)	
How many counties in New York State?
```{r}
print(nrow(county.df))
```

### Question 1(d)	
What’s the 3-digit fips code of Broome County?
```{r}
Broome <- county.df[county.df$NAME10=="Broome",]
print(Broome$fips)
```

### Question 1(e)	
Compute descriptive statistics of the population column (P0010001), including total, minimum, maximum, mean, median, and skewness. 
```{r, message=FALSE, warning=FALSE}
print(Broome$P0010001)
### descriptive statistics: 
pop <- county.df$P0010001
if(!require(fBasics)) {install.packages("fBasics")}
library(fBasics)
stat <- basicStats(pop)
stat
stat["Sum",1]
stat["Minimum",1]
stat["Maximum",1]
stat["Mean",1]
stat["Median",1]
stat["Skewness",1]
```

### Question 1(f)	
Create a histogram and a boxplot of the population.
```{r}
hist(pop,breaks = 20)
boxplot(pop)
```


## Question 2
Use R package UScensus2010tract to complete the following tasks:    (20 pt.)
```{r}

```

### Question 2(a)	
Plot a map of New York census tracts using the plot function.


### Question 2(b)
Compute the total population based on census tracts.
```{r}
library(UScensus2010tract)
data("new_york.tract10")
tract <- new_york.tract10
plot(tract)
tract.df <- tract@data
tract.df
print(paste("Total populations is: ", sum(tract.df$P0010001)))
```

### Question 2(c)
Select all census tracts in Broome County and plot the map. 
```{r}
Broome <- tract[tract.df$county=="007",]
plot(Broome)
```

### Question 2(d)
What’s the total population of Broome County?
```{r}
df <- Broome@data
nrow(df)
pop <- df$P0010001
sum(pop)
```

### Question 2(e)
Create a histogram and a boxplot of population based on census tracts of Broome County.
```{r}
hist(pop)
boxplot(pop)
```

### Question 2(f)
Select the first five columns of the shapefile of Broome County census tract; add a new population ratio column (= census tract population / county population); save the new shapefile to the hard drive. 
```{r, message=FALSE, warning=FALSE}
Broome <- Broome[,1:5]
df <- Broome@data
df$pop.pct <- df$P0010001/sum(pop)
df
Broome@data <- df
library(rgdal)
writeOGR(Broome,dsn = "output",layer = "Broome",driver = "ESRI Shapefile",overwrite_layer = TRUE)
```


## Question 3

Use R packages raster and ncdf4 to complete the following tasks:     (20 pt.)


### Question 3(a)		
Load the multi-band raster dataset “NDVI.nc” into R.
```{r}
library(raster)
library(ncdf4)
ndvi <- brick("NDVI.nc")
ndvi
```

### Question 3(b)		
Get the basic information about the dataset, including the number of rows, columns, cells, and bands; spatial resolution, extent, bounding box, and projection.
```{r}
nrow(ndvi)
ncol(ndvi)
ncell(ndvi)
nlayers(ndvi)
res(ndvi)
extent(ndvi)
bbox(ndvi)
projection(ndvi)
```

### Question 3(c)	
Aggregate all bands to generate a mean NDVI raster and a maximum NDVI raster; save the two new raster datasets to the hard drive. 
```{r}
ndvi.mean <- mean(ndvi)
plot(ndvi.mean)

ndvi.max <- max(ndvi)
plot(ndvi.max)

ndvi.range <- range(ndvi)
plot(ndvi.range)

writeRaster(ndvi.mean,filename = "output/ndvi.mean.tif",overwrite=TRUE)
writeRaster(ndvi.max,filename = "output/ndvi.max.tif",overwrite=TRUE)
```

### Question 3(d)			
Plot the maps of monthly NDVI of the year 2001
```{r}
#plot(ndvi)
plot(ndvi,13:24)
```

### Question 3(e)	
Create histograms of monthly NDVI of the year 2001.
```{r, message=FALSE, warning=FALSE}
ndvi2001 <- subset(ndvi,13:24)
ndvi2001
plot(ndvi2001)
hist(ndvi2001)
```

### Question 3(f)			
Plot the NDVI map of July 2000; click any location with data on the map and retrieve the NDVI time series for all years; plot the NDVI time series of the selected location. 
```{r, message=FALSE, warning=FALSE}
plot(ndvi,7)
# values <- click(ndvi,n=1,xy=FALSE)
# plot(as.vector(values),type="b")

values <- ndvi[50,50]
plot(as.vector(values), type="b")
```

